Predicting Remaining Useful Life of NASA Aircraft Engines

### 2025-02-25 index.qmd
Updated data visualization with the status column and added the univariate data

Though CPH is typically used with health data, we wanted to determine
how CPH can be used in other fields. As such, we will study the
remaining useful life of a NASA aircraft engine using data provided by
NASA for the study on propagation modeling [@saxena2008]. In particular,
we will be examining their data from what they have deemed engine two.
From the literature review, we have noted that we will need to be
concerned with the constant hazards ratio within our data and determine
how we will accommodate censored data when creating our model.

@@

@@

test <- read.table(testurl, header = TRUE, sep = '')
head(test)
```

@@

train <- read.table(trainurl, header = TRUE, sep = '')
head(test)
```

@@

```{r}

>>>>

Though CPH is typically used with health data, we wanted to determine
how CPH can be used in other fields. As such, we will study the
remaining useful life (RUL) of a NASA aircraft engine using data
provided by NASA for the study on propagation modeling [@saxena2008]. In
particular, we will be examining their data from what they have deemed
engine two. From the literature review, we have noted that we will need
to be concerned with the constant hazards ratio within our data and
determine how we will accommodate censored data when creating our model.

@@

For both test and train datasets, we will need to add a column
indicating status. We will use 0 for still working and 1 for the last
cycle when the engine fails. For the test data, we know that none of the
engines have failed so all values in the test column will be 0 for still
working. For the training data, we know that the data includes all
cycles up to the last cycle when the machine will fail. As such, we will
have 0s in all cycles except the last cycle of the machine.

@@

test <- read.table(testurl, header = TRUE, sep = '') %>%
  mutate(status = 0)
head(test, n = 5)
```

@@

train <- read.table(trainurl, header = TRUE, sep = '') %>%
  group_by(id) %>%
  mutate(status = if_else(row_number() == n(), 1, 0)) %>%
  ungroup()

head(train, n = 5)
```

@@

From the study the data is from, we know that some of the sensors will
not play a role in the RUL. As such, we will run a univariate test of
all covariates to determine which ones will actually impact the RUL. We
will use the p-value provided by the Wald test with a significance level
of $\alpha = 0.05$.

```{r}
covariates <- c("os1", "os2", "os3", "sm1", "sm2", "sm3", "sm4", "sm5", "sm6", "sm7",
                "sm8", "sm9", "sm10", "sm11", "sm12", "sm13", "sm14", "sm15", "sm16",
                "sm17", "sm18", "sm19", "sm20", "sm21")

univ_formulas <- sapply(covariates,
                        function(x) as.formula(paste('Surv(time, status)~', x)))

univ_models <- lapply( univ_formulas, function(x){coxph(x, data = train)})
# Extract data
univ_results <- lapply(univ_models,
                       function(x){
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (",
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test",
                                        "p.value")
                          return(res)
                         })
as.data.frame(univ_results)
```
